---
title: "Decathlon Analysis"
output: html_notebook
---
# Introduction

The data used for this analysis contained results from the decathalon at the Olympics and Decaster. 
Their results for each event was recorded along with their final points total.
I assumed the same name across both competitions is the same athlete



# Packages needed and load in data

```{r, warning=FALSE}
library(tidyverse)
library(here)

clean_decathlon_data <- read_csv(here("clean_data/clean_decathlon_data.csv"))
```

# Who had the longest long jump seen in the data?

```{r}
clean_decathlon_data%>%
  select(name, event, result) %>%
  filter(event == "long_jump")%>%
  slice_max(result)
```

# What was the average 100m time in each competition?

```{r}
clean_decathlon_data %>%
  filter(event == "100m") %>%
  group_by(competition) %>%
  summarise(avg_100m_time = mean(result))
```
# Who had the highest total points across both competitions?

```{r}
clean_decathlon_data %>%
  group_by(name)%>%
  summarise(total_points = sum(points))%>%
  slice_max(total_points)
```

# What was the shot-put scores for the top three competitors in each competition?

```{r}
clean_decathlon_data%>%
  select(name, competition, event, result) %>%
  filter(event == "shot_put") %>%
  group_by(competition) %>%
  slice_max(result, n=3)
```

# What was the average points for competitors who ran the 400m in less than 50 seconds vs. those that ran 400m in more than 50 seconds?

```{r}
clean_decathlon_data %>%
  filter(event == "400m") %>%
  mutate(`400m_time` = if_else(result < 50, "under 50", "over 50")) %>%
  group_by(`400m_time`) %>%
  summarise(avg_points = mean(points))
 
```

