---
title: "R Notebook"
output: html_notebook
---
# Packages needed and load in data
```{r}
library(tidyverse)
library(here)

decathlon_data <- read_csv(here("clean_data/clean_decathlon_data.csv"))
```
# Who had the longest long jump seen in the data?

```{r}
decathlon_data%>%
  filter(event == "long_jump")%>%
  slice_max(result)
```

#What was the average 100m time in each competition?

```{r}
decathlon_data %>%
  filter(event == "100m") %>%
  group_by(competition) %>%
  summarise(avg_time = mean(result))
```
Who had the highest total points across both competitions?

```{r}
decathlon_data %>%
  group_by(name)%>%
  summarise(total_time = sum(points))%>%
  slice_max(total_time)
```

#What was the shot-put scores for the top three competitors in each competition?

```{r}
decathlon_data%>%
  select(name, competition, event, result) %>%
  filter(event == "shot_put") %>%
  group_by(competition) %>%
  slice_max(result, n=3)
```

#What was the average points for competitors who ran the 400m in less than 50 seconds vs. those that ran 400m in more than 50 seconds?

```{r}
decathlon_data %>%
  filter(event == "400m") %>%
  mutate(`400m_time` = if_else(result < 50, "under 50", "over 50")) %>%
  group_by(`400m_time`) %>%
  summarise(avg_points = mean(points))
 
```

