---
title: "Halloween Candy Analysis"
output:
  html_document:
    df_print: paged
---

# Introduction

The data used for the analysis below came from three datasets (2015,
2016, 2017). The datasets for each year contained ratings of various items,
some questions about the person and random questions which were unnecessary for this analysis.  
The three raw datasets were all cleaned then combined to enable analysis
to be carried out.
Summary of cleaning stages :

* Cleaned all column names
* Removed any irrelevant columns
* Pivotted some of the data to make for easier analysis
* Renamed various candy and country inputs to make it uniform
* Filtered ages outwith 4-90 and turned them to NA's
* Removed any items which weren't candies
* Made sure all datasets had the same columns to enable binding to take place



Assumptions made whilst cleaning :

1. Removed the coloured M&Ms as this was a hidden proxy question, as stated on the website containing the datasets.
2. A lot of renaming was done for the countries to account for misspellings and 'creative' country naming. For example:
    * Canada - "soviet canuckistan" and "canae"
    * United States - "cascadia" and "trumpistan"
    * UK - all the different countries that make up the UK and "god's country"

Things I could have done with more time when cleaning to improve results :

* Rescued some NA countries in the 2017 dataset by checking if there is an input in the state/region column
* Rescued some NA ages by checking if they've been entered in the country column 



##### Load in packages and data
```{r message=FALSE}
library(tidyverse)
library(here)

candy_data <- read_csv(here("clean_data/clean_all_candy.csv"))
glimpse(candy_data)
```


# Analysis

#### What is the total number of candy ratings given across the three years.
(Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
sum(!is.na(candy_data$rating))
```


#### What was the average age of people who are going out trick or treating?

```{r}
candy_data%>%
  filter(going_trick_or_treating == "Yes") %>%
  summarise(avg_age = mean(age, na.rm = TRUE))
  
```

#### What was the average age of people who are not going trick or treating?

```{r}
candy_data%>%
  filter(going_trick_or_treating == "No") %>%
  summarise(avg_age = mean(age, na.rm = TRUE))
```

#### For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r, message=FALSE}
candy_data %>%
  filter( !is.na(rating))%>%
  group_by(rating, candy)%>%
  summarise(count = n()) %>%
  slice_max(count) 

```

#### How many people rated Starburst as despair?

```{r}
candy_data %>%
  filter(candy == "starburst" & rating == "despair")%>%
  summarise(starburst_despair_ratings = n())
```

#### *For the next three questions, count despair as -1, joy as +1, and meh as 0.*

*Made a new table and only kept the joy and despair ratings (meh = 0 so
not needed in calculations). Changed "joy" to "1" and "despair" to "-1"*

```{r}
rating_keep <- c("joy", "despair")

candy_rating_filtered <- candy_data%>%
  filter(rating %in% rating_keep)%>%
  mutate(rating = if_else(rating == "joy", 1, -1))

```

*For the answers below, I included the top two most popular for each category, as
full sized candy bar was the most popular for every category.*

#### What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r, message=FALSE}
gender_filter <- c("Male", "Female", "Other")

candy_rating_filtered  %>%
  filter( gender %in% gender_filter)%>%
  group_by(gender, candy)%>%
  summarise(points = sum(rating))%>%
  slice_max(points, n=2)
```

#### What was the most popular candy bar in each year?

```{r, message=FALSE}
candy_rating_filtered%>%
  group_by(year, candy)%>%
  summarise(points = sum(rating))%>%
  slice_max(points, n=2)

```

#### What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r, message=FALSE}
candy_rating_filtered%>%
  group_by(country, candy)%>%
  summarise(points = sum(rating))%>%
  slice_max(points, n=2)
```


# Conclusion

Any full sized candy bar was the most popular, this is most likely down to the vague description hence covering a wide range of candy. If necessary, further analysis could be done on the data provided, a useful analysis could be comparing the the popularity change of each candy year-on-year.

 To make future surveys more conducive to better analysis, I would suggest using the same survey questions and format. Another way of improving results could be to introduce a drop down menu with answers for questions like age and country, this would eliminate any misspellings.